Model-agnostic Interpretable Machine Learning

### 2021-07-29 2021-bips-practice-phd-defence/2021-bips-practice-phd-defence.Rmd
structure talk

- A Brief History of Interpretability
- A Closer Look at Model-agnostic interpretability
- methods:
  - Feature Effects: PDP, ICE and ALE
  - Feature Importance: PFI
- Overview of Stuff I did in my PhD 
- Measuring Model Complexity
- Conditional Feature Importance Using Subgroups
- Inference of Effects

@@

@@

Special focus on old, established tools, and make them fit for

>>>>

- A Brief History of Interpretable Modeling
- A Closer Look at Model-agnostic interpretability
- methods:
  - Feature Effects: PDP, ICE and ALE
  - Feature Importance: PFI
- Overview of Stuff I did in my PhD 
- Measuring Model Complexity
- Conditional Feature Importance Using Subgroups
- Inference of Effects

# Before You Pin Me Down

To get this out of the way, here is how I use "Interpretable Machine Learning":

It's an umbrella term for approaches that analyze or describe prediction mechanisms in models that were learned on data.

# A Brief History of Interpretable Modeling

Interpretable Models with data have a longer history, but the field of interpretable machine learning is still very young as consolidated field.

* 1800s: Linear regression models
* 1960: Rule-based ML (AID, CHAID, CART, association rules, ...)
* 2001: Random Forest
* 2021: Deep Learning Boom kicked off
* 2016: Interpretable ML / Explainable AI boom

While 200 years of history, consolidation as larger field only started in 2016.

TODO: Add citation report and or google trends image

# Why Interpretability?

* The more non-linear and complex a model, the more difficult to understand how it operates
* Trust
* Debugging
* Control 
* ...

# Interpretable Models

* GAM and other linear models
* Trees and decision rules
* More complex models, but interpretation requires more work or post-processing

# Model-agnostic Interpretation

Basic idea: Treat the model as a black box, not allowed to look inside.

Idea very powerful: Allows modularity, works for most models, results are comparable across models

TODO: CITE SIPA paper here

# PDP and ICE

# PFI

TODO: Explain how PFI works

# PFI and PDP similarity

CITE paper GC here

# Difference between model-agnostic and model-specific interpetation

* Analyzing components vs. analyzing behavior
* Components not very powerful, unless they directly describe behavior
* For complex model, components do not directly tell us about behavior of model
* Behavior more comparable

@@

TODO: Draw image with Ipad

@@

Special focus on old, established tools, and make them fit for right use in practice

Also some more specific papers

* subgroups
* inference
* quantifying

and some contributions to

* MOC
* relative feature importance
* SIPA
* feature importance and PDP (GC)

# Paper: Subgroups

## Problem: Dependent features

Show example with dependent features for PFI

## Soluation: Conditional variants

Show solution

* But: Not meaningful for PDP, or other measure more useful (LOESS)
* But: Conditioning unclear
* But: Intepretation difficult

## Our solution: Permute in subgroups

TODO: Show algorithm

## Some results from our paper

TODO: Show results or application

# Paper: IML Inference

## Problem: Uncertainty quantification

PDP and PFI and most other just descriptive tools
Often practictioner draw confidence intervals
but only based on monte carlo approximation

* But what is even the ground truth?
* And what about the uncertainty in the model fit?
* How could we estimate the uncertainty?

## Solution: Treat as statistical estimator

## How to Compute the variance

## Problem of too low variance estimates

# Paper: Quantifying

TODO: Copy as much as possible from the talk that already exists for this paper

Motivation: want to quantify in a model-agnostic way how complex a model is

## Functional Decomposition

Explain how functional decomposition works

## ALE

Formula:

TODO: For estimation show the images

## The three measures

* Mean effect complexity
* non linearity
* number of features

## Using in benchmarking (multi-objective)

## Thank You for Your Attention
